Relationships in the extremes and their influence on competition and coexistence

### 2021-07-16 paper/Paper.Rnw
fleshed out example

For clarification, we explain the how storage effects operate in maintaining biodiversity and the conditions that are required for it to do so.
We emphasize that although this paper focuses on storage effects, we do consider other mechanisms of modern coexistence theory, though we quickly realize the importance of ATAs is solely for storage effects; hence, it is still a paper generally about the importance of ATAs for coexistence because. 
Storage effects allow for species to fluctuate over time while escaping extinction. 
In order for storage effect to be a functioning mechanism in coexistence, populations must fluctuate differently in response to environmental fluctuations, there must be a relationship between envrionmental responses and competition, and there must variation in responses at different life history stages. 
In a simple two species system, differences in environmental response allow for one species to grow to high densities while the other is very low. 
Competition is positively related to environmental response for the high density species (species A), stabalizing its growth. 
Conversely, competition is more negatively related to environmental response for the low density species (species B), leading to large fluctuations in population growth rate in both directions. 
Seemingly, this may appear like a scenario for extiction of species B, but it can, in fact, be at an advantage due to differences in life history stages. 
Different responses at different life stages buffer the species against extinction by limiting negative growth rates, but not positive ones. 
Examples of life stages that can buffering against negative growth rates are insensitive adults or dormant seeds that persist even when environmental response amoung others in the population is low (Chesson et al 2012). 
Thus, higher fluctuation in growth rate can lead to a higher long term mean growth rate. 
In this paper, we investigate the consequences of having ATAs in the relationship between fluctuations environmental response and competitive pressure. More specifically, we answer: Does having higher association in a tail end of the EC relationship, while maintaining overall association, change the contribution of storage effect?

[As we know], anthropogenic climate change threatens almost all aspects of life including via extreme climatic events, which have only become more extreme - increasing in magnitude and frequency - and have contributed to both the increase in global means as well as an increase in environmental variability. 
There are many studies on the influence of environmental extremes and variability on competition and coexistence (Parepa et al 2013, Diez et al 2012); however, the study of the influence of relationships and interactions between drivers in their extremes on competition and coexistence, is lacking in the scientific literature. 
Interacting environmental drivers can be asymmetrically related in the extreme values (as shown in Gosh et al 2020 AER). Thus we have implemented our tools for the study ATAs to explore the importance of environmental relationships in their extremes on competition and coexistence. 
Environmental variability has been classically shown to promote coexistence via storage mechanisms in a lottery model of competition (Chesson and Warner 1981, Warner and Chesson 1985), hence we continue/move forward with this simple model to examine the effect of extreme relationships between environmental variables on the contribution of storage effects to coexistence.

The lottery model within the framework of coexistence theory provides a suitable opportunity to explore extreme relationships and the consequences to coexistence.

(establish criteria for storage effects and establish connection to extreme events to introduce example)
Results main ideas + example

@@

final parahraph:goals
%Jasmin, I think the below makes a pretty good start at tryinbg to nail down a good way to express our goals
%in the last para of the Intro, so please incorporate these into your last para. If you have a better way
%to phrase these, please put both so we can compare at some point when we meet.
G1) To augment the modern theory of coexistence (MTC) to show how asymmetric tail association (ATA)
plays a role. The MTC decomposes a measure of coexistence into components due to each of several mechanisms. 
We address how ATA contributes to these components. We here examine mechanisms of coexistence that apply in 
models exhibiting solely temporal variation; spatial variation can be considered in future work.
G2) To explore, using a simple model, the relative magnitudes of the contributions to coexistence that come
from ATA compared to other mechanisms.

>>>>

For clarification, we explain the how storage effects operate in maintaining biodiversity and the conditions that are required for it to do so.
We emphasize that although this paper focuses on storage effects, we do consider other mechanisms of modern coexistence theory, though we quickly realize the importance of ATAs is solely for storage effects; hence, it is still a paper generally about the importance of ATAs for coexistence because. 
Storage effects allow for species to fluctuate over time while escaping extinction. 
In order for storage effect to be a functioning mechanism in coexistence, populations must fluctuate differently in response to environmental fluctuations, there must be a relationship between envrionmental responses and competition, and there must variation in responses at different life history stages. 
In a simple two species system, differences in environmental response allow for one species to grow to high densities while the other is very low. 
Competition is positively related to environmental response for the high density species (species A), stabalizing its growth. 
Conversely, competition is more negatively related to environmental response for the low density species (species B), leading to large fluctuations in population growth rate in both directions. 
Seemingly, this may appear like a scenario for extiction of species B, but it can, in fact, be at an advantage due to differences in life history stages. 
Different responses at different life stages buffer the species against extinction by limiting negative growth rates, but not positive ones. 
Examples of life stages that can buffering against negative growth rates are insensitive adults or dormant seeds that persist even when environmental response amoung others in the population is low (Chesson et al 2012); environment and competiton do not effect the buffered life stages.  
Thus, higher fluctuation in growth rate can lead to a higher long term mean growth rate. 
In this paper, we investigate the consequences of having ATAs in the relationship between fluctuations environmental response and competitive pressure. More specifically, we answer: Does having higher association in a tail end of the EC relationship, while maintaining overall association, change the contribution of storage effect?

[As we know], anthropogenic climate change threatens almost all aspects of life including via extreme climatic events, which have only become more extreme - increasing in magnitude and frequency - and have contributed to both the increase in global means as well as an increase in environmental variability (citation). 
There are many studies on the influence of environmental extremes and variability on competition and coexistence (Parepa et al 2013, Diez et al 2012); however, the study of the influence of relationships and interactions between drivers in their extremes on competition and coexistence, is lacking in the scientific literature. 
Interacting environmental drivers can be asymmetrically related in the extreme values (as shown in Gosh et al 2020 AER). 
Environmental drivers directly influence competitive pressure experienced by populations through environmental response; competition is always a function of environmental response (Chesson et al 2012, Ellner et al 2016).
Thus we have implemented our tools for the study of ATAs to explore the importance of environmental relationships in their extremes on competition and coexistence. 
Environmental variability has been classically shown to promote coexistence via storage mechanisms in a temporal lottery model of competition (Chesson and Warner 1981, Warner and Chesson 1985), hence we continue/move forward with this simple model to examine the effect of extreme relationships between environmental responses on the contribution of storage effects to coexistence. Moreover, the lottery model within the framework of coexistence theory provides a suitable opportunity to explore relationships in their extremes and the consequences to coexistence.

To flesh out the idea of how ATAs in environmental response might manifest and modfiy the outcome of coexistence, we introduce an example of a two species grass-forb plant community (inspired by Chesson et al 2012). The community follows the criteria required for temporal storage effects to operate. The forb dominates and has high fecundity when a sufficient threhold of rain is reached in the early season. The grass dominates and has high fecundity when there is sufficeint rain in the late season. Competition is influenced by the resource comsumption of growing juvenile plants. Dormant seeds are unaffected by competiton and unfavorable environment and can persist in the community. When rain conditions fluctuate from year to year, relative abundance and dominance of the grass-forb community also fluctuate year to year. This produces a relationship between the grass's and forb's environmental responses (fecundity) that is symmetrically associated. ATAs in environmental response can be caused by extreme climatic events. If the community is affected by frequent and long drought with some years of normalcy, the grass-forb environmental response relationship may become more correlated in the lower values or left tail associated. Conversely, if the community is effected by frequent and long floods with some years of normalcy, the grass-forb environmental relationship may become more correlated in the higher values or right-tail associated. In Figure 2 we see a possible outcome of ATAs between environmental responses. In simulating their populations,  ATAs can cause species to have longer periods of extreme dominance and lower growth rates overall.

@@

final parahraph:goals
%Jasmin, I think the below makes a pretty good start at tryinbg to nail down a good way to express our goals
%in the last para of the Intro, so please incorporate these into your last para. If you have a better way
%to phrase these, please put both so we can compare at some point when we meet.
G1) To augment the modern theory of coexistence (MTC) to show how asymmetric tail association (ATA)
plays a role. The MTC decomposes a measure of coexistence into components due to each of several mechanisms. 
We address how ATA contributes to these components. We here examine mechanisms of coexistence that apply in 
models exhibiting solely temporal variation; spatial variation can be considered in future work.
G2) To explore, using a simple model, the relative magnitudes of the contributions to coexistence that come
from ATAs compared to other mechanisms.
